State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-06-09 ltx/BrainIDE.tex
switching to different sty for journal part 2
made some little typo changes. Also what's going on with the document class?

\author{Dean~R.~Freestone,~\IEEEmembership{Graduate Student Member,~IEEE,}
        Parham~Aram,~\IEEEmembership{Graduate Student Member,~IEEE,}
        Michael~Dewar}
        Kenneth~Scerri,~\IEEEmembership{Member,~IEEE,}
		David~B.~Grayden,~\IEEEmembership{Member,~IEEE,}
        Visakan~Kadirkamanathan,~\IEEEmembership{Member,~IEEE,}% <-this % stops a space% <-this % stops a space

@@

>>>>

\author{Dean~R.~Freestone,~\IEEEmembership{Graduate Student Member,~IEEE,}
        Parham~Aram,~\IEEEmembership{Graduate Student Member,~IEEE,}
        Michael~Dewar,
        Kenneth~Scerri,~\IEEEmembership{Member,~IEEE,}
		David~B.~Grayden,~\IEEEmembership{Member,~IEEE,}
        Visakan~Kadirkamanathan,~\IEEEmembership{Member,~IEEE,}}% <-this % stops a space% <-this % stops a space

@@

\begin{pmatrix}\alpha & \beta\\
	\gamma & \delta
\end{pmatrix}

### 2010-06-09 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
made some little typo changes. Also what's going on with the document class?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[twocolumn,11pt,a4paper]{article}		%<-use this for checking equation length
% \documentclass[12pt]{iopart}

@@

\title[State and Parameter Estimation for Spatio-Temporal Neural Fields]{State and Parameter Estimation for Spatio-Temporal Neural Fields}

\author{Dean R. Freestone$^{1,5}$, Parham Aram$^2$, Michael Dewar$^3$, Kenneth Scerri$^4$, David B. Grayden$^{1,5}$, and Visakan Kadirkamanathan$^2$}

@@

 &&T_s \int_\Omega {\boldsymbol{\phi} (\mathbf{r}) \int_\Omega {\boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}') \mathbf{x}_t ) d\mathbf{r}'}d\mathbf{r}}\boldsymbol{\theta} \nonumber\\

@@

\int_\Omega{\boldsymbol{\phi} (\mathbf{r}) e_t(\mathbf{r})d\mathbf{r}}. 

@@

	 \lefteqn{\mathbf{x}_{t+1} \nonumber = T_s\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r}) \int_\Omega {\boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta}} \nonumber\\

@@

	\lefteqn{\mathbf{x}_{t+1} \nonumber= \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t}\nonumber \\
& +& \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}(\mathbf{r})e_t(\mathbf{r})d\mathbf{r}}.

@@

where the superscripts $m$ and $c$ stand for mean and covariance. Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are  used to correct the predicted states. The state estimates from the forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the cross-covariance matrix of the states, $\mathbf M$, needs to be computed. A summary of the URTSS procedure is given in Algorithm~\ref{UKFAlgorithm}. It should be noted that  the disturbance, $e_t\left(\mathbf{r}\right)$ and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTS is used, otherwise the state vector should be augmented with the noise terms. The additive disturbance form reduces the dimension and the number of sigma points used in the algorithm.

@@

Although the system is nonlinear, the parameters of the system are linear with respect to the state. This is exploited by our procedure where the parameter estimation uses a least squares (LS) method that minimises the sum of the squared errors (of a predicted state update) with each new state vector sequence estimate (see~\ref{App:LeastSquaresAppendix}).

@@

   		\subfigure[][]{\includegraphics{./Graph/ExperimentFigureEEGPy.pdf}} \\

@@

\begin{figure*}
	\label{fig:FreqAnalysis} 
	\begin{center}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/BasisChar.pdf}}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/SensorChar.pdf}}
\subfigure[][]{\includegraphics[width=1.8in]{./Graph/ObservationFrequencyResponse.pdf}}\\
	\end{center}
	\caption{Frequency analysis of the. \textbf{a.} The spatial frequency response of a 1-dimensional neural field basis functions project. The symmetry of the field basis functions, connectivity kernel basis functions and sensors allow for the 1-dimensional representation of the spectral characteristics. The dashed line shows the cut-off frequency (-3 dB point). \textbf{b} The spatial frequency response of a 1-dimensional model sensor. The dashed line shows the cut-off frequency (-3 dB point), indicating a wider bandwidth than the field basis function. \textbf{c.} Two 2-dimensional spatial frequency (in power dB, averaged over 400 time steps) of the observed field.} 
\end{figure*}
Following this, the distance between basis function, $\Delta_{\phi}$, was set to $2.5$ and the basis function width, $\sigma_{\phi}^2$, was set to 2.5, 2.63 mm full width at half maximum. \mike{Analysis said 0.44 from Parham's derivation, but 2.5 still overlaps and allows more high frequency details to be represented.}\parham{maybe we can argue that this width is for perfect recunstruction up to 0.19 Hz which requires the smaller distance between basis functions and hence more number of basis functions but we choose to go for lower number of basis function and loosing high frequency contents.}

\subsection{State and Parameter Estimates} 
\label{sec:state_and_param_results}
To demonstrate the performance of the parameter and state estimation, $100$ realisations of the data were generated. Each realisation consisted of 500~ms of data and the estimation was applied to the final 400~ms, allowing the model to stabilise.\dean{transient!} The initial state and parameters were unknown to the estimator. The distribution of the parameters and the resulting kernel are shown in figure~\ref{fig:Parameters} and figure~\ref{fig:KernelEstimates}, respectively. Although a wide distribution can be seen over the parameter estimates, the ratio between them remains within a narrow range as shown in figure~\ref{fig:ParametersRatio}. The state estimates for one realisation were evaluated by comparing the field reconstruction to the true field using the mean over time of the Root Mean Square Error (RMSE) and standard deviation of the RMSE giving $0.5\pm 0.02$. The RMSE of the field estimates over time for 100 realizations is also shown in figure~\ref{fig:RMSE}. A snapshot of the true field, the estimated field and their corresponding error is illustrated in figure~\ref{fig:FieldEstimate}. The convergence of the parameter estimation is shown in figure~\ref{fig:ParametersConvergence}.

@@

 |\boldsymbol\Phi(\boldsymbol\nu_{c})|^2=\frac{1}{2}|\boldsymbol\Phi(\mathbf 0)|^2

@@

 \sigma_{\nu}^2=\frac{2\boldsymbol\nu_{c}^\top \boldsymbol\nu_{c}}{\ln 2 }.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[twocolumn,10pt,a4paper]{article}		%<-use this for checking equation length
%\documentclass[12pt]{iopart}

@@

\title{State and Parameter Estimation for Spatio-Temporal Neural Fields}

\author{Dean R. Freestone$^{1,5}$, Parham Aram$^2$, Michael Dewar$^3$, Kenneth Scerri$^4$, David B. Grayden$^{1,5}$,\\ and Visakan Kadirkamanathan$^2$}

@@

 &&T_s \int_\Omega {\boldsymbol{\phi} (\mathbf{r}) \boldsymbol{\theta}^{\top} \int_\Omega {\boldsymbol{\psi} \left(\mathbf{r}-\mathbf{r}'\right) f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right) \mathbf{x}_t \right)d\mathbf{r}'}d\mathbf{r}} \nonumber\\

@@

\int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) e_t\left(\mathbf{r}\right)d\mathbf{r}}. 

@@

	 \lefteqn{\mathbf{x}_{t+1} \nonumber = T_s\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r}) \int_\Omega {f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta}} \nonumber\\

@@

	\mathbf{x}_{t+1} = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t + \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}(\mathbf{r})e_t(\mathbf{r})d\mathbf{r}}.

@@

where the superscripts $m$ and $c$ stand for mean and covariance. Since the observation equation is linear (equation~\ref{ObservationEquation}) the standard Kalman Filter update equations are  used to correct the predicted states. The state estimates from the forward filtering are used to form a new set of sigma points for the smoother, as described above. To compute the smoother gain, the predicted cross-covariance matrix of the states is used with the covariance matrix. A summary of the URTSS procedure is given in Algorithm~\ref{UKFAlgorithm}. It should be noted that  the disturbance, $e_t\left(\mathbf{r}\right)$ and the measurement noise, $\boldsymbol{\varepsilon}_t$, are additive. Therefore, the additive form of the URTS is used, otherwise the state vector should be augmented with the noise terms. The additive disturbance form reduces the dimension and the number of sigma points used in the algorithm.

@@

Although the system is nonlinear, the parameters of the system are linear with respect to the state. This is exploited by our procedure where the parameter estimation uses a least squares (LS) method that minimises the sum of the squared errors (of a predicted state update) with each new state vector sequence estimate (see~\ref{LeastSquaresAppendix}).

@@

   		\subfigure[][]{\includegraphics[width=0.3\textwidth]{./Graph/ExperimentFigureEEGPy.pdf}} \\

@@

\begin{figure*}
	\label{fig:FreqAnalysis} 
	\begin{center}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/BasisChar.pdf}}
\subfigure[][]{\includegraphics[scale=0.8]{./Graph/SensorChar.pdf}}
\subfigure[][]{\includegraphics[width=1.8in]{./Graph/ObservationFrequencyResponse.pdf}}\\
	\end{center}
	\caption{Frequency analysis of the. \textbf{a.} The spatial frequency response of a 1-dimensional neural field basis functions project. The symmetry of the field basis functions, connectivity kernel basis functions and sensors allow for the 1-dimensional representation of the spectral characteristics. The dashed line shows the cut-off frequency (-3 dB point). \textbf{b} The spatial frequency response of a 1-dimensional model sensor. The dashed line shows the cut-off frequency (-3 dB point), indicating a wider bandwidth than the field basis function. \textbf{c.} Two 2-dimensional spatial frequency (in power dB, averaged over 400 time steps) of the observed field.} 
\end{figure*}
Following this, the distance between basis function, $\Delta_{\phi}$, was set to $2.5$ and the basis function width, $\sigma_{\phi}^2$, was set to 2.5, 2.63 mm full width at half maximum. \mike{Analysis said 0.44 from Parham's derivation, but 2.5 still overlaps and allows more high frequency details to be represented.}

\subsection{State and Parameter Estimates} 
\label{sec:state_and_param_results}
To demonstrate the performance of the parameter and state estimation, $100$ realisations of the data were generated. Each realisation consisted of 500~ms of data and the estimation was applied to the final 400~ms, allowing the model to stabilise.\dean{transient!} The initial state and parameters were unknown to the estimator. The distribution of the parameters and the resulting kernel are shown in figure~\ref{fig:Parameters} and figure~\ref{fig:KernelEstimates}, respectively. Although a wide distribution can be seen over the parameter estimates, the ratio between them remains within a narrow range as shown in figure~\ref{fig:ParametersRatio}. The state estimates for one realisation were evaluated by comparing the field reconstruction to the true field using the mean over time of the Root Mean Square Error (RMSE) and standard deviation of the RMSE giving $0.5\pm 0.02$. The RMSE of the field estimates over time for 100 realizations is also shown in figure~\ref{fig:RMSE}. A snapshot of the true field, the estimated field and their corresponding error is illustrated in Figure~\ref{fig:FieldEstimate}.

@@

 |\boldsymbol\Phi(\boldsymbol\nu_c)|^2=\frac{1}{2}|\boldsymbol\Phi(\mathbf 0)|^2

@@

 \sigma_{\nu}^2=\frac{2\boldsymbol\nu_c^\top \boldsymbol\nu_c}{\ln 2 }.
